---
title: "hw3"
author: "Joe Despres"
date: "11/13/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---


## HW3: Power Analysis

Due Tr Nov 19th in D2L. One file per sumbission (please sumbit either doc, pdf or html from R-markdown) showing the code, results and your interpretation of results, all in the same file.

### 1. Power analysis: main effects

**Research Goal**:  to estimate the effect of Body Mass Index (BMI) on Systolic Blood Pressure (SBP), while accounting by sex differences.

Sex differences can take the form of a main effect of sex on SBP or an interaction between sex and BMI.


Baseline model:  SBP~BMI+SEX

Design: a total of N subjects (N/2 men and N/2 women) will be recruited, for each subject SBP and BMI will be measured.

**Task**: to produce power curves (power versus sample size) by effect size (see below).

**Simulation setting**.
  * Equal proportion in the sample of male and female.
  * Simulate BMI assuming `p(BMI|Sex=Female)=normal(mean=26.5,variance=30)` and `p(BMI|Sex=Male)=normal(mean=27.4,variance=16.7)`. 
  * Simulate your response using `SBP=mu+M*b1+Z*b2+error` where mu=120, M is a dummy variable for Male,  b1=-3, Z =BMI-mean(BMI) (note since there are equal numbers of male and female, mean(BMI) is just the average BMI of male and female).
  * Aassume an error variance of 300.
	
**Estimation**: For estimation purposes use `lm()` to regress SBP on Sex and centered BMI. Consider rejection consider a Type-I error rate of 0.05.

**1.1.** Use Monte Carlo Methods (at least 5000 replicates) to estimate power curves (power versus sample size, N=30,50,100,200,500,1000,3000) by effect size (b2=0,0.2, 0.3 and 0.5 ).

```{r}
sim <- function(n, b2, nreps, b1 = -3, mu = 120){
  
REJ_RATE <- matrix(ncol = length(b2), nrow = length(n))
colnames(REJ_RATE) <- b2
rownames(REJ_RATE) <- n
reject <- rep(NA, nreps)
  
for(i in 1:length(n)){
 for(j in 1:length(b2)){
   for(k in 1:nreps){

bmim <- rnorm(n = n[i]/2,  mean=27.4, sd=sqrt(16.7))
bmif <- rnorm(n = n[i]/2,  mean=26.5, sd=sqrt(30))

data <- cbind(c(rep(1, n[i] / 2), rep(0, n[i] / 2)), c(bmim, bmif))
data <- as.data.frame(data)

colnames(data) <- c("M", "BMI")

data$Z   <- data$BMI - mean(data$BMI)
error    <- rnorm(n[i], mean = 0, sd = sqrt(300))
data$SBP <- mu + (data$M * b1) + (data$Z * b2[j]) + error

fit <- lm(SBP ~ M + Z, data)

reject[k] <- summary(fit)$coef[3, 4] < 0.05

}
REJ_RATE_[i, j] <- mean(reject)
  }
    }
cbind.data.frame(n, REJ_RATE_)
}

```



```{r}
set.seed(55555)
n <- c(30,50,100,200,500,1000,3000)
b2 <- c(0,0.2, 0.3,0.5)
powers <- sim(n, b2, nreps = 1e4)
REJ_RATE - powers[1:4,]
beepr::beep()
```

```{r}
plot(0,ylim=c(0,1),xlim=range(1e4),ylab='Rejection rate',xlab='Sample size') abline(h=seq(from=0,to=1,by=.1),col=8,lwd=.5) 
abline(v=seq(from=0,to=max(N),by=500),lwd=.5,col=8)
for(i in 1:ncol(REJ_RATE)){
lines(x=N,y=REJ_RATE[,i],col=i) }
abline(h=.05,col=4,lty=2)
```

**1.2.** What is the minimum sample size needed to achieve a power of at least 80% if the effect size is 0.3?

## First Approach
Let's first look at a graph of the powers vs sample size. 

```{r}
plot(powers$n, powers$`0.3`, col = "red", xlab = "Sample", ylab = "Power", pch = 19)
abline(h = 0.8, lty = "dotted")
abline(v = 1000, lty = "dotted")
```
Here we can see that a sample of 1000 brings us fairly close. However, the increase in power associated with an increase in sample diminishes quite rapidly. Let us guess approximate from the chart that the sample size should be between 1100 and 1150

```{r}
n <- seq.int(from = 1100,to = 1200, by = 2)
powers3 <- sim(n, b2 = 0.3, nreps = 1e4)

plot(powers3$n ,powers3$`0.3`, 
     xlab = "Sample Size", ylab = "Power", 
     main = "Sample Size of 1118 obtains a power of 0.8")
lines(powers3$n, powers3$`0.3`)
abline(h = 0.8, lty = "dotted")
abline(v = 1124)
abline(v = 1140)
```
Here we can see that to obtain a power above 0.80 we need a sample size above 1110. To be absoultely sure we will have a power above 0.8 we probally want a sample bigger than 1150


## Second Approach

This way we will set a while loop to run until the power reaches 0.8 and then print the sample size.

```{r}
sim3 <- function(nreps, from, to,  b1 = -3, b2 = 0.3, mu = 120){
i <- 2
n <- seq.int(from = from, to = to, by = 2)
  
REJ_RATE <- matrix(ncol = length(b2), nrow = length(n))
colnames(REJ_RATE) <- "power"
rownames(REJ_RATE) <- n
reject <- rep(NA, nreps)
condition <- 0
while(condition < 0.8){
  sample <- i
  i <- i + 2
  for(k in 1:nreps){
    
    bmim <- rnorm(n = n[i]/2,  mean=27.4, sd=sqrt(16.7))
    bmif <- rnorm(n = n[i]/2,  mean=26.5, sd=sqrt(30))
    
    data <- cbind(c(rep(1, n[i] / 2), rep(0, n[i] / 2)), c(bmim, bmif))
    data <- as.data.frame(data)
    
    colnames(data) <- c("M", "BMI")
    
    data$Z   <- data$BMI - mean(data$BMI)
    error    <- rnorm(n[i], mean = 0, sd = sqrt(300))
    data$SBP <- mu + (data$M * b1) + (data$Z * b2) + error
    
    fit <- lm(SBP ~ M + Z, data)
    reject[k] <- summary(fit)$coef[3, 4] < 0.05
  }
 condition <- mean(reject)
 }
print(paste("To obtain a Power of", condition,"Requires a sample size of", n[sample]))
}

```

```{r}
sim3(nreps = 1e4, from = 1000, to = 1500, b1 = -3, b2 = 0.3, mu = 120)
```
I think the problem with this way opposed to the other way is that you will get the very first .8. This may be better than the fi However, These numbers are quite close, we expect them to be slightly different. therefore in order to obtain a power of 0.8 we require a sample of between 1120 and 1135. 


### 2. Power Analysis: interaction effects

Assume that the effect of BMI on SBP is higher (0.4) in male than in female (0.2). Develop a MC study to estimate the power to detect this interaction as a function of sample size (go up to N=10,000). Suggestion: extend the baseline model by adding a sex by BMI interaction and assess the power to detect an interaction of the size above-specified.

```{r}
sim2 <- function(n, nreps, mu = 120, b1 = 0.2, b2 = 0.2){

REJ_RATE <- matrix(ncol = 1, nrow = length(n))
colnames(REJ_RATE) <- "Power"
reject <- rep(NA, nreps)
  
for(i in 1:length(n)){
   for(k in 1:nreps){

bmim <- rnorm(n = n[i]/2,  mean=27.4, sd=sqrt(16.7))
bmif <- rnorm(n = n[i]/2,  mean=26.5, sd=sqrt(30))

data <- cbind(c(rep(1, n[i] / 2), rep(0, n[i] / 2)), c(bmim, bmif))
data <- as.data.frame(data)

colnames(data) <- c("M", "BMI")

data$Z <- data$BMI - mean(data$BMI)
error  <- rnorm(n[i], mean = 0, sd = sqrt(300))

b1 <- -3
b2 <- ifelse(data$M == 1, 0.4, 0.2)

data$SBP <- mu + (b1 * data$M) + (b2 * data$Z) + error

fit <- lm(SBP ~  M + Z + M * Z, data)

reject[k] <- summary(fit)$coef[4, 4] < 0.05

}
REJ_RATE[i,] <- mean(reject)
 }
cbind.data.frame(n, REJ_RATE)
      }
```

```{r}
n <- c(30, 50, 100, 200, 500, 1000, 3000, 5000, 10000)

powers5 <- sim2(n, nreps = 1e4)
powers5
```
Yes we can get a power of over 50% with a sample less than 10,000.


